### 2015-07-03 narrative/intro.tex
some small changes -- a work in progress....

{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
Richard Hamming's remark, made over 50 years ago at the dawning of the age of
large scale scientific computation, is even more relevant today as we prepare
for scientific simulation at the exascale. Hamming was a pioneer in the fields
of coding theory, scientific computing and information theory who would
instantly sympathize with the central objective of this proposal--minimizing
the time to knowledge.

Avoiding impending bottlenecks in exascale scientific discovery will require
new research into managing and storing the large amounts of data that will be
produced by simulations and analyzed for months afterwards.
%
In this project we will demonstrate a cooperative approach for storing data
where the user and the storage system work together to achieve the best
possible data quality and performance given data features, storage tier
characteristics, and current system state. We will investigate novel techniques
to facilitate efficient mapping of data objects, even partitioning individual
variables, from the user space onto multiple storage tiers and enable
application-guided data reductions/transformations to address capacity and
bandwidth bottlenecks while constraining the error to be within user provided
bounds.
% 
We will address the associated I/O and storage challenges in the context of
current and emerging storage landscapes and expedite insights into critical
scientific processes, while demonstrating the validity of our approach in key
DOE domains. We will research techniques for creating a Scalable Storage
Software Infrastructure, integrating services from the middleware layer with
Storage System features to support the novel strategy of distributing data both
horizontally and vertically across the storage system.  The negotiation between
these layers will be provided as a fundamental service, allowing users to
easily save ``the best'' amount of data, automatically placed throughout the
entire storage stack.

The metric we are most interested in optimizing is time to knowledge.  Current
approaches to addressing the I/O bottleneck fall into two broad categories:
parallel file system approaches that optimize the throughput for an entire
system, and I/O middleware approaches that optimize the performance of a single
application. Both approaches have been successful to date, but are unlikely to
overcome the remaining obstacles in reaching exascale. Instead of trying to
optimize throughput, we will seek to reduce the time to knowledge. This is the
most significant metric for scientific applications where the desired outcome
is not storing data, but rather executing a knowledge extraction process on the
data. Moreover, we aim to perform the optimization not for a single
application, but rather for workloads within a multi-user multi-application
system environment.

Our approach targets the expected characteristics of exascale storage hardware.
The storage layer will be partitioned into multiple heterogeneous tiers with
vastly different performance characteristics. The layers will be further
differentiated by the constraints on capacity and data lifetime within each
layer. Tape archival storage will still maintain data long term, but access to
this data will be orders of magnitude slower than the next layer and naively
accessing data from archival storage will greatly impact productivity.
%
Our approach in this proposal is to enable researchers to place more knowledge
into the SSIO system so that data can be recast from a bunch of bytes to
information which the system can then place and optimize the time to gain
knowledge from exascale simulations.  We first motivate this by a problem our
team has been working on in the past six months, and then recast this problem
into a series of research questions and challenges which must be addressed in
order to optimize the SSIO system for exascale science.

@@

This solution worked well for the first day of run but after the second day more problems occurred: the variability of the time it took to write the data made the I/O prohibitively costly, because even though the project had reserved the majority of resources at the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds. There was no method for the application or the user to monitor and predict this resource usage and to dynamically change to write less data when the I/O time was going to be longer. The greatest problem, however, comes after the simulation is going to be completed: even to visualize the reduced data, just to get a first look at the visualization data, it could take over a month because some of the timesteps will be on tape and other
timesteps will be on the file system. There is no method to just look at the data quickly at some level of quality quickly and understand which regions of the simulation to focus on in more detail.

>>>>

{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
Richard Hamming's remark, made over 50 years ago at the dawning of the age of
large scale scientific computation, is even more relevant today as we prepare
for scientific simulation at exascale. 
The central objective of this proposal, ``minimizing the time knowledge" from 
scientific computing, aligns perfectly this sentiment. 
%Hamming was a pioneer in the fields of coding theory, scientific computing and 
%information theory who would instantly sympathize with the central objective of 
%this proposal--minimizing the time to knowledge.

Avoiding impending bottlenecks in exascale scientific discovery will require
new research into managing and storing the large amounts of data that will be
produced by simulations and analyzed for months afterwards.
%
In this project we will demonstrate a cooperative approach for storing data
where the user and the storage system work together to achieve the best
possible utility and performance given data features, storage tier
characteristics, and current system state. We will investigate novel techniques
to facilitate efficient mapping of data objects, even partitioning individual
variables, from the user space onto multiple storage tiers and enable
application-guided data reductions/transformations to address capacity and
bandwidth bottlenecks, while constraining errors to be within user provided
bounds.
%

We will address the associated I/O and storage challenges in the context of
current and emerging storage landscapes and expedite insights into critical
scientific processes, while demonstrating the validity of our approach in key
DOE domains. We will research techniques for creating and using a Scalable 
Storage Software Infrastructure, and integrating services from the middleware 
layer with Storage System capabilities to support novel strategies for distributing 
data both horizontally and vertically across the storage system.  Negotiation between
these layers will be provided as a fundamental service, allowing users to
easily save ``the best'' amount of data, which is automatically placed throughout the
entire storage stack.

The metric we are most interested in optimizing is time to knowledge.  Current
approaches to addressing the I/O bottleneck fall into two broad categories:
parallel file system approaches that optimize the throughput for an entire
system, and I/O middleware approaches that optimize the performance of a single
application. Both approaches have been successful to date, but are unlikely to
overcome the remaining obstacles in reaching exascale. 
Instead of trying to optimize throughput, we will seek to reduce the time to knowledge. 
This is the most significant metric for scientific applications where the desired outcome
is not storing data, but rather executing a knowledge creation and extraction processes 
on the data. 
Moreover, we aim to perform the optimization not for a single
application, but rather for workloads within a multi-user multi-application
system environment.

Our approach targets the expected characteristics of exascale storage hardware.
The storage layer will be partitioned into multiple heterogeneous tiers with
vastly different performance characteristics. The layers will be further
differentiated by the constraints on capacity and data lifetime within each
layer. Tape archival storage will still maintain data long term, but access to
this data will be orders of magnitude slower than the next layer and naively
accessing data from archival storage will greatly impact productivity.
%
Our approach in this proposal is to enable researchers to place more knowledge
into the SSIO system so that data can be recast from a bunch of bytes to
information, which the system can then place and optimize the time to gain
knowledge from exascale simulations.  We first motivate this by a problem our
team has been working on in the past six months, and then recast this problem
into a series of research questions and challenges which must be addressed in
order to optimize the SSIO system for exascale science.

@@

This solution worked well for the first day of run but after the second day more problems occurred: the variability in the time that it took to write the data made the I/O prohibitively expensive, because even though the project had reserved the majority of the resources at the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds. There was no method for the application or the user to monitor and predict this resource usage and to dynamically adapt to write less data when the I/O time was going to be longer. The greatest problem, however, comes after the simulation   completes: even to visualize the reduced data, just to get a first look at the visualization data, it could take over a month because some of the timesteps will be on tape and other timesteps will be on the file system. There is no method to just look at the data quickly at some level of quality quickly and understand which regions of the simulation to focus on in more detail.
